---
title: "ma615 midterm slide"
author: "Luyin Fu"
date: "10/20/2019"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
pacman::p_load("magrittr","knitr","ggplot2","tidyr", "dplyr", "kableExtra", "esquisse")
```


## Introduction

In this report, we look at 
- world urbanization
- world carbon dioxide emission
- relationship between these two variables


```{r warning=FALSE}
data <- readxl::read_xls("/Users/tsuyu/Downloads/Data\ science\ in\ R/project/World\ Bank\ Data/climate_change.xls",
         sheet = "Data")
pop <- data %>% filter(`Series code`=="SP.POP.TOTL")
urbanpop <- data %>% filter(`Series code`=="SP.URB.TOTL")

Country_info <- readxl::read_xls("/Users/tsuyu/Downloads/Data\ science\ in\ R/project/World\ Bank\ Data/climate_change.xls", sheet="Country")

```

## World urban population

```{r warning=FALSE}
pop1 <-  dplyr::select(pop, `Country name`, `2005`, `1995`)
pop1 %<>% rename(pop2005 = `2005`, pop1995=`1995`)
urbanpop1 <- urbanpop %>% 
  dplyr::select(`Country name`, `2005`, `1995`) %>%
  rename(urbanpop2005 = `2005`, urbanpop1995=`1995`)
urban_pop_perc <- merge(pop1, urbanpop1)  
urban_pop_perc %<>% 
  mutate(perc2005=as.numeric(urbanpop2005)/as.numeric(pop2005),
         perc1995=as.numeric(urbanpop1995)/as.numeric(pop2005),
         perc_incre=perc2005-perc1995) 

region <- Country_info %>% 
  filter(`Capital city`!="..") %>%
  select(`Country code`, `Country name`, Region)

urban_pop_perc <- merge(region, urban_pop_perc)
urban_pop_perc %<>% drop_na()
```

In this section, we look specifically at 1995 and 2005

```{r}
ggplot(urban_pop_perc)+
  geom_point(aes(x=perc1995, y=perc2005, color=Region))+
  labs(x ="urban population percentage 1995", 
       y = "urban population percentage 2005")+
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9", "#F0E442", "#6600FF", "#CCFF00", "#CC79A7"))
```



## Comparing 


```{r}
urban_pop_perc1 <- urban_pop_perc %>% 
  select(`Country name`, perc2005, Region) %>%
  mutate(year="2005") %>%
  rename(perc=perc2005)
urban_pop_perc2 <- urban_pop_perc %>% 
  select(`Country name`, perc1995, Region) %>%
  mutate(year="1995")%>%
  rename(perc=perc1995)
urban_pop_perc3 <- rbind(urban_pop_perc1,urban_pop_perc2)

ggplot(urban_pop_perc3)+aes(x=year, y=perc, fill=Region)+
  geom_boxplot()+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9", "#F0E442", "#9999CC", "#CCFF00", "#CC79A7"))
```

```{r}
ECA <- urban_pop_perc3 %>% 
  filter(Region=="Europe & Central Asia") %>%
  group_by(year)%>%
  arrange(perc)%>%
  slice(1)
MENA <- urban_pop_perc3 %>% 
  filter(Region=="Middle East & North Africa") %>%
  group_by(year)%>%
  arrange(perc)%>%
  slice(1)
SSA <- urban_pop_perc3 %>% 
  filter(Region=="Sub-Saharan Africa") %>%
  group_by(year)%>%
  arrange(desc(perc))%>%
  slice(1)
```

<span style="color: blue>The outliner for Europe & Central Asia Region in 1995 and 2005 is `r ECA[[1,1]]`.
<br>
In Middle East & North Africa, the country with an abnormally low urban population percentage is `r MENA[[1,1]]`
<br>
the African country that has an abnormally high urban population percentage in both years is `r SSA[[1,1]]`.</span>

## Countries

```{r}
urban_pop_top_05 <- urban_pop_perc %>% 
  dplyr::select(`Country name`, perc2005) %>%
  arrange(desc(perc2005))%>%
  slice(1:10)
urban_pop_top_05$perc2005 <- round(urban_pop_top_05$perc2005,4)
urban_pop_top_95 <- urban_pop_perc %>% 
  dplyr::select(`Country name`, perc1995) %>%
  arrange(desc(perc1995))%>%
  slice(1:10)
urban_pop_top_95$perc1995 <- round(urban_pop_top_95$perc1995,4)
urban_pop_top <- cbind(urban_pop_top_95, urban_pop_top_05)
colnames(urban_pop_top) <- rep(c('Country', 'Urban Population %'),2)
kable(urban_pop_top, format = "html", booktabs=TRUE,
      caption = "Countries with Lagest Urban Population Percentage")%>%
  add_header_above(c("1995"=2, "2005"=2))
```


## top

```{r}
urban_pop_perc_10 <- urban_pop_perc %>% 
  dplyr::select(`Country name`, perc1995, perc2005, perc_incre) %>%
  arrange(desc(perc_incre)) %>%
  slice(1:10)
urban_pop_perc_10$perc_incre <- round(urban_pop_perc_10$perc_incre, 4)
urban_pop_perc_10$perc1995 <- round(urban_pop_perc_10$perc1995, 4)
urban_pop_perc_10$perc2005 <- round(urban_pop_perc_10$perc2005, 4)
colnames(urban_pop_perc_10) <-c("Country", "Urban Population % 1995", 
                                "Urban Population % 2005", "Increment")
kable(urban_pop_perc_10, format = "latex", booktabs=TRUE, 
      caption = "Countries with bigest Increase 1995-2005") %>%
  kable_styling(latex_options = "HOLD_position") 
```


## Carbon Dioxide Emission Per Capita

Carbon dioxide emissions are those stemming from the burning of fossil fuels and the manufacture of cement. They include carbon dioxide produced during consumption of solid, liquid, and gas fuels and gas flaring. Data are in million metric tons.

```{r}
CO2_emiss <- data %>% filter(`Series code`=="EN.ATM.CO2E.PC")
CO2_emiss1 <- CO2_emiss %>% select(`Country name`, `2005`, `1995`)%>%
  rename(emission05 = `2005`, emission95=`1995`)
CO2_emiss1 <- merge(CO2_emiss1, region) %>% 
  filter(emission05 != '..' & emission95!= '..' )


C1 <- CO2_emiss1 %>% 
  select(`Country name`, emission05, Region) %>%
  mutate(year="2005") %>%
  rename(emission=emission05)
C2 <- CO2_emiss1 %>% 
  select(`Country name`, emission95, Region) %>%
  mutate(year="1995") %>%
  rename(emission=emission95)
CO2_emiss2<- rbind(C1, C2)
CO2_emiss2$emission <- as.numeric(CO2_emiss2$emission)

ggplot(CO2_emiss2)+aes(x=year, y=emission, fill=Region)+
  geom_boxplot()+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9", "#F0E442", "#9999CC", "#CCFF00", "#CC79A7"))

```


## Urbanization and Carbon Dioxide Emission

```{r}
Data1 <- merge(CO2_emiss2, urban_pop_perc3)

ggplot(Data1)+
  geom_point(aes(x=perc, y=emission, color=Region))+
  facet_grid(year~.)+
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9", "#F0E442", "#6600FF", "#CCFF00", "#CC79A7"))+
  labs(title=" ",
        x ="urban population percentage", 
       y = "CO2 emission (million metric tons)")
```


## looking for linear relationship(1995)

```{r}
Data3 <- Data1 %>% filter(year=="1995")
ggplot(Data3)+aes(x=perc, y=emission, color=Region)+
  geom_point()+
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9", "#F0E442", "#6600FF", "#CCFF00", "#CC79A7"))+
  geom_smooth(method = lm)+
  labs(title="1995",
        x ="urban population percentage", 
       y = "CO2 emission (million metric tons)")+
    facet_wrap(~Region)

```


## 2005

```{r}
Data2 <- Data1 %>% filter(year=="2005")
ggplot(Data2)+aes(x=perc, y=emission, color=Region)+
  geom_point()+
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9", "#F0E442", "#6600FF", "#CCFF00", "#CC79A7"))+
  geom_smooth(method = lm)+
  labs(title="2005",
        x ="urban population percentage", 
       y = "CO2 emission (million metric tons)")+
    facet_wrap(~Region)

```





